---
title: "Non-Target Screening Demo"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Non-Target Screening Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Goal

This article shows a complete non-target screening workflow built with the
current streamfind project framework. The example uses
`ProjectNonTargetAnalysis`, stores project state in DuckDB, and relies on demo
mass-spectrometry files distributed through `streamfindData`.

Compared with the legacy engine-based examples, the recommended pattern is now:

1. Open a project-backed object.
2. Import raw files into the project database.
3. Attach sample metadata such as replicates and blanks.
4. Define a workflow with `Method_NonTargetAnalysis_*()` constructors.
5. Run the workflow and inspect persistent results from R or the app.

# Requirements

The examples below assume that `streamfindData` is installed in addition to
`streamfind`.

```{r, eval = FALSE}
install.packages("streamfindData")
```

The workflow in this article uses the wastewater TOF demo files and the
reference table of spiked chemicals with MS2 information shipped by
`streamfindData`.

# Load packages

```{r, eval = FALSE}
library(streamfind)
library(streamfindData)
library(data.table)
library(plotly)
```

# Get demo resources

First, collect the demo mzML files. Here the example keeps the wastewater
analyses used in the development workflow.

```{r, eval = FALSE}
ms_files <- streamfindData::get_ms_file_paths()
ms_files <- ms_files[grepl("ww_", ms_files)]

ms_files
```

Next, load the reference chemicals table and split it into internal standards
and general suspects. The exact columns used by the NTA methods are kept small
on purpose.

```{r, eval = FALSE}
reference_db <- streamfindData::get_ms_tof_spiked_chemicals_with_ms2()

reference_db <- data.table::as.data.table(reference_db)[
  ,
  .(name, formula, mass, rt, fragments, tag)
]

internal_standards <- reference_db[grepl("IS", tag) & !is.na(rt)]
suspects <- reference_db[!grepl("IS", tag)]
```

If you want to include MetFrag or transformation-product assignment, prepare
their input tables separately. streamfind also provides template exporters for
these reference tables.

# Create the project

The project object persists imported analyses, workflow configuration, results,
and cache objects inside a DuckDB database.

```{r, eval = FALSE}
project_db <- file.path(tempdir(), "streamfind_nta_demo.duckdb")
project_id <- "wastewater_demo"

if (file.exists(project_db)) file.remove(project_db)

nta <- open_ProjectNonTargetAnalysis(
  db = project_db,
  project_id = project_id,
  file_paths = ms_files
)
```

Because the project is persistent, you can later reopen it with only the
database path and `project_id`.

```{r, eval = FALSE}
nta <- open_ProjectNonTargetAnalysis(
  db = project_db,
  project_id = project_id
)
```

# Add replicate and blank information

Replicate and blank assignments are project metadata. They are used later by
workflow steps such as blank subtraction and matrix suppression handling.

```{r, eval = FALSE}
nta$set_replicate_names(c(
  rep("neg_blank", 3),
  rep("pos_blank", 3),
  rep("neg_influent", 3),
  rep("pos_influent", 3),
  rep("neg_effluent", 3),
  rep("pos_effluent", 3)
))

nta$set_blank_names(c(
  rep("neg_blank", 3),
  rep("pos_blank", 3),
  rep("neg_blank", 3),
  rep("pos_blank", 3),
  rep("neg_blank", 3),
  rep("pos_blank", 3)
))
```

You can inspect the project at any point:

```{r, eval = FALSE}
print(nta)
info(nta)
nta$get_domain()
nta$list_tables()
```

# Inspect available processing methods

Project classes expose a method registry so you can confirm which processing
steps are available for the domain.

```{r, eval = FALSE}
registry <- nta$available_processing_methods()
names(registry)

projects_overview()$ProjectNonTargetAnalysis$processing_methods
```

# Build a complete workflow

The recommended workflow is built from method objects and wrapped in a
`Workflow`. The example below mirrors a realistic NTS sequence:

- feature finding
- MS1 and MS2 loading
- component creation and annotation
- internal-standard matching
- feature grouping and filling
- blank subtraction and feature filtering

```{r, eval = FALSE}
workflow <- Workflow(list(
  Method_NonTargetAnalysis_FindFeatures(
    rtWindows = data.frame(rtmin = numeric(), rtmax = numeric()),
    ppmThreshold = 10,
    noiseThreshold = 250,
    minSNR = 3,
    minTraces = 3L,
    baselineWindow = 200,
    maxWidth = 250,
    baseQuantile = 0.99,
    debugAnalysis = "",
    debugMZ = 0,
    debugSpecIdx = -1L
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS1(
    filtered = FALSE,
    rtWindow = c(-1, 1),
    mzWindow = c(-1, 6),
    minTracesIntensity = 250,
    mzClust = 0.008,
    presence = 0.5
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS2(
    filtered = FALSE,
    minTracesIntensity = 10,
    isolationWindow = 1.3,
    mzClust = 0.008,
    presence = 0.5
  ),
  Method_NonTargetAnalysis_CreateComponents(
    rtWindow = c(-5, 5),
    minCorrelation = 0.8,
    debugRT = 0,
    debugAnalysis = ""
  ),
  Method_NonTargetAnalysis_AnnotateComponents(
    maxIsotopes = 8L,
    maxCharge = 1L,
    maxGaps = 1L,
    ppm = 10,
    debugComponent = "",
    debugAnalysis = ""
  ),
  Method_NonTargetAnalysis_FindInternalStandards(
    suspects = internal_standards,
    ppm = 10,
    sec = 15,
    ppmMS2 = 10,
    mzrMS2 = 0.008,
    minCosineSimilarity = 0.7,
    minSharedFragments = 3L,
    filtered = TRUE
  ),
  Method_NonTargetAnalysis_FilterInternalStandards(
    idLevels = c(1L, 2L, 3L)
  ),
  Method_NonTargetAnalysis_GroupFeatures(
    method = "internal_standards",
    rtDeviation = 5,
    ppm = 10,
    minSamples = 1L,
    binSize = 5,
    debug = FALSE,
    debugRT = 0
  ),
  Method_NonTargetAnalysis_FillFeatures(
    withinReplicate = FALSE,
    filtered = FALSE,
    rtExpand = 10,
    mzExpand = 0.01,
    maxPeakWidth = 30,
    minTracesIntensity = 1000,
    minNumberTraces = 5L,
    minIntensity = 5000,
    rtApexDeviation = 5,
    minSignalToNoiseRatio = 3,
    minGaussianFit = 0.2,
    debugFG = ""
  ),
  Method_NonTargetAnalysis_BlankSubtraction(
    blankThreshold = 5,
    rtExpand = 10,
    mzExpand = 0.005
  ),
  Method_NonTargetAnalysis_FilterFeatures(
    minIntensity = 10000,
    removeIsotopes = TRUE,
    removeAdducts = TRUE,
    removeLosses = TRUE
  )
))
```

Attach the workflow to the project:

```{r, eval = FALSE}
set_workflow(nta, workflow)
show(nta$get_workflow())
```

# Optional extensions

The project workflow can be extended with suspect screening, matrix suppression
correction, MetFrag screening, or transformation-product assignment.

## Suspect screening

```{r, eval = FALSE}
workflow_extended <- Workflow(c(unclass(workflow), list(
  Method_NonTargetAnalysis_SuspectScreening(
    suspects = suspects,
    ppm = 5,
    sec = 10,
    ppmMS2 = 10,
    mzrMS2 = 0.008,
    minCosineSimilarity = 0.7,
    minSharedFragments = 3L,
    filtered = TRUE
  )
)))
```

## Matrix suppression correction

```{r, eval = FALSE}
matrix_suppression_step <- Method_NonTargetAnalysis_CorrectMatrixSuppression(
  refBlankReplicate = NA_character_,
  mpRtWindow = 10
)
```

## MetFrag screening

```{r, eval = FALSE}
metfrag_step <- Method_NonTargetAnalysis_MetFragScreening(
  metfrag_path = "C:/Tools/MetFragCommandLine-2.5.0.jar",
  database_type = "LocalCSV",
  database_path = "transformation_products_template.csv",
  ppm = 10,
  sec = 15,
  ppmMS2 = 10,
  mzrMS2 = 0.008,
  top_n = 5L,
  score_types = "FragmenterScore",
  score_weights = 1,
  pre_processing_candidate_filter = c("UnconnectedCompoundFilter", "IsotopeFilter"),
  post_processing_candidate_filter = "InChIKeyFilter",
  maximum_tree_depth = 2L,
  number_threads = 1L,
  use_smiles = TRUE,
  filtered = FALSE,
  java_path = "java",
  run_dir = "",
  debug = FALSE,
  extra_params = list()
)
```

# Run the workflow

```{r, eval = FALSE}
nta$run_workflow()
```

The results are written back into the project database. This means you can
close the R session, reopen the project later, and continue from the stored
state.

# Inspect the results

## Retrieve tables

```{r, eval = FALSE}
features_all <- get_features(nta)
features_filtered <- get_features(nta, filtered = TRUE)

internal_standards_found <- get_internal_standards(nta)
suspects_found <- get_suspects(nta)
```

When transformation-product assignment is part of the workflow:

```{r, eval = FALSE}
transformation_products_found <- get_transformation_products(nta)
```

## Filter by a target mass or reference row

The project getters accept either scalar mass windows or full reference rows.

```{r, eval = FALSE}
features_subset <- get_features(
  nta,
  mass = internal_standards[4, ],
  ppm = 20,
  sec = 60,
  filtered = TRUE
)
```

## TIC overview

```{r, eval = FALSE}
plot_spectra_tic(
  nta,
  downsize = 1,
  levels = 1,
  groupBy = "replicate"
)
```

## Feature counts

```{r, eval = FALSE}
plot_features_count(nta, groupBy = "replicate")
```

## Feature profiles

```{r, eval = FALSE}
plot_features_profile(
  nta,
  mass = internal_standards[4, ],
  ppm = 20,
  sec = 60,
  corrected = TRUE,
  filtered = TRUE
)
```

## Feature overview plots

```{r, eval = FALSE}
plot_features(
  nta,
  mass = internal_standards[4, ],
  ppm = 20,
  sec = 60,
  filtered = TRUE
)
```

## MS1 and MS2 inspection

```{r, eval = FALSE}
plot_features_ms1(
  nta,
  analyses = 11,
  mass = internal_standards[4:6, ],
  ppm = 20,
  sec = 60,
  filtered = FALSE
)

plot_features_ms2(
  nta,
  analyses = 11,
  mass = internal_standards[4:6, ],
  ppm = 20,
  sec = 60,
  filtered = FALSE
)
```

## Matrix suppression overview

When matrix suppression correction is included in the workflow, you can inspect
the computed suppression data directly.

```{r, eval = FALSE}
matrix_suppression <- get_matrix_suppression(
  nta,
  rtWindowVal = 10
)

plot_matrix_suppression(
  nta,
  rtWindowVal = 10,
  colorBy = "replicates"
)
```

This can be combined with the TIC plot in a subplot for a quick quality check.

```{r, eval = FALSE}
tic_plot <- plot_spectra_tic(
  nta,
  downsize = 1,
  levels = 1,
  groupBy = "replicate"
)

matrix_plot <- plot_matrix_suppression(
  nta,
  colorBy = "replicates"
)

plotly::subplot(
  tic_plot,
  matrix_plot,
  nrows = 2,
  shareX = TRUE,
  titleY = TRUE
)
```

# Reopen the same project later

Because the analysis state is stored in DuckDB, reopening the project is
lightweight:

```{r, eval = FALSE}
nta <- open_ProjectNonTargetAnalysis(
  db = project_db,
  project_id = project_id
)
```

This is especially useful when the full workflow takes time and you only want
to continue inspecting results.

# Open the app

The same project can be launched in the app for manual review.

```{r, eval = FALSE}
nta$run_app()
```

# Summary

This demo illustrates the full project-based NTS pattern in streamfind:

1. Use `streamfindData` to get example files and reference tables.
2. Open a `ProjectNonTargetAnalysis` backed by DuckDB.
3. Store replicate and blank metadata in the project.
4. Build a workflow from `Method_NonTargetAnalysis_*()` steps.
5. Run the workflow once and reuse the saved results later.
6. Inspect features, TICs, MS1/MS2, matrix suppression, suspects, and internal
   standards from the same project object.

For additional interactive review, open the project in the app rather than
rebuilding the workflow each time.
