---
title: "Report Peak Finding Merck V3"
subtitle: "Integration in StreamFind and Structural Based Screening"
author: "Ricardo Cunha"
date: "`r format(Sys.time(), '%d %B, %Y')`"
toc: true
format:
  html:
    grid:
      sidebar-width: 300px
      body-width: 1200px
      margin-width: 300px
      gutter-width: 1.5rem
    other-links:
      - text: StreamFind
        href: https://odea-project.github.io/StreamFind/
        icon: github
    html-math-method: katex
    theme: journal
    code-tools: true
    anchor-sections: true
    code-fold: false
    code-overflow: scroll
    code-line-numbers: false
    code-copy: true
    cap-location: top
    self-contained: true
    embed-resources: true
    toc: true
    toc-location: left
execute:
  warning: false
---

```{r}
#| label: setup
#| include: false

options(DT.options = list(
  dom = "lfrtip",
  autoWidth = TRUE,
  pageLength = 10,
  lengthMenu = c(10, 20, 50, 75, 100),
  initComplete = htmlwidgets::JS(
    "function(settings, json) {",
    "$(this.api().table().body()).css({'font-size': '12px'});",
    "$(this.api().table().body()).css({'white-space': 'nowrap'});",
    "}"
  )
))

simple_table_opt <- list(
  dom = "t",
  paging = FALSE,
  searching = FALSE,
  info = FALSE
)

library(knitr)
library(kableExtra)
library(data.table)
library(DT)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)

# devtools::load_all()

resources_file <- file.path(
  getwd(),
  "dev",
  "merck_peak_finding",
  "dev_resources.R"
)

if (!file.exists(resources_file)) {
  resources_file <- file.path(getwd(), "dev_resources.R")
}

project_dir <- file.path(dirname(resources_file), "data")
dir.create(project_dir, recursive = TRUE, showWarnings = FALSE)

source(resources_file)
```

# Integration in StreamFind

## Project Setup

```{r}
#| label: merck-cleanup
#| include: false

# if (file.exists(project_db)) file.remove(project_db)
```

```{r}
#| label: merck-engine-setup

example_key <- "ACC1_28203"
project_id <- paste0("merck_peak_finding_", example_key)
project_db <- file.path(project_dir, paste0(project_id, ".duckdb"))

ex_targets <- merck2_db_ex[example %in% example_key]
ex_files <- merck2_files_ex$file_path[grepl(example_key, merck2_files_ex$example)]

nta <- open_ProjectNonTargetAnalysis(
  db = project_db,
  project_id = project_id,
  file_paths = ex_files
)

nta$set_replicate_names(c("blank", "sample"))
nta$set_blank_names(c("blank", "blank"))
```

```{r}
#| label: merck-inspect-engine

nta
```

```{r}
#| label: merck-inspect-tic

#plot_spectra_tic(nta, downsize = 3, levels = 1)
```

## Workflow

```{r}
#| label: merck-workflow

workflow <- Workflow(list(
  Method_NonTargetAnalysis_FindFeatures(
    rtWindows = data.frame(rtmin = 300, rtmax = 3400),
    ppmThreshold = 15,
    noiseThreshold = 15,
    minSNR = 3,
    minTraces = 3L,
    baselineWindow = 200,
    maxWidth = 250,
    baseQuantile = 0.99,
    debugAnalysis = "",
    debugMZ = 0,
    debugSpecIdx = -1L
  ),
  Method_NonTargetAnalysis_CreateComponents(
    rtWindow = c(-5, 5),
    minCorrelation = 0.5,
    debugRT = 0,
    debugAnalysis = ""
  ),
  Method_NonTargetAnalysis_AnnotateComponents(
    maxIsotopes = 8L,
    maxCharge = 1L,
    maxGaps = 1L,
    ppm = 10,
    debugComponent = "",
    debugAnalysis = ""
  ),
  Method_NonTargetAnalysis_BlankSubtraction(
    blankThreshold = 5,
    rtExpand = 20,
    mzExpand = 0.005
  ),
  Method_NonTargetAnalysis_FilterFeatures(
    removeIsotopes = TRUE,
    removeAdducts = TRUE,
    removeLosses = TRUE
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS1(
    rtWindow = c(-1, 1),
    mzWindow = c(-1, 6),
    mzClust = 0.008,
    presence = 0.5,
    minTracesIntensity = 50,
    filtered = FALSE
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS2(
    isolationWindow = 1.3,
    mzClust = 0.008,
    presence = 0.5,
    minTracesIntensity = 10,
    filtered = FALSE
  )
))

set_workflow(nta, workflow)
```

```{r}
#| label: merck-inspect-workflow

show(get_workflow(nta))
```

```{r}
#| label: merck-run-workflow
#| results: "hide"

run_workflow(nta)
```

## Results

```{r}
#| label: merck-inspect-features-count
#| fig-height: 7
#| fig-width: 10

p1 <- plot_features_count(nta, filtered = TRUE, groupBy = "replicate") %>%
  plotly::layout(annotations = list(
    list(
      x = 0.2, y = 1, xref = "paper", yref = "paper",
      text = "Before Filtration", showarrow = FALSE,
      font = list(size = 16)
    )
  ))

p2 <- plot_features_count(nta, filtered = FALSE, groupBy = "replicate") %>%
  plotly::layout(annotations = list(
    list(
      x = 0.2, y = 1, xref = "paper", yref = "paper",
      text = "After Filtration", showarrow = FALSE,
      font = list(size = 16)
    )
  ))

plotly::subplot(
  p1, p2,
  nrows = 1,
  margin = 0.05,
  shareY = TRUE,
  titleX = TRUE,
  titleY = TRUE
)
```

### Targets

```{r}
#| label: merck-inspect-results
#| fig-height: 10
#| fig-width: 10

plot_features(
  nta,
  analyses = 2,
  mass = ex_targets,
  ppm = 20,
  sec = 60,
  showDetails = TRUE
)
```

```{r}
#| label: merck-inspect-results-2

fts <- get_features(
  nta,
  analyses = 2,
  mass = ex_targets,
  ppm = 20,
  sec = 120,
  filtered = TRUE
)

colnames(fts)

ex_targets[!ex_targets$name %in% fts$id, ]
```

### Components

```{r}
#| label: merck-inspect-components

map_features(
  nta,
  analyses = 2,
  groupBy = "feature_component",
  rt = ex_targets[7, ]$rt,
  sec = 60,
  globalNormalization = TRUE,
  showDetails = TRUE
)
```

# App

```{r}
#| label: run_app
#| eval: false

nta <- open_ProjectNonTargetAnalysis(
  db = project_db,
  project_id = project_id
)

nta$run_app()
```

# Identification

```{r}
#| label: merck-suspects-screening



rcpp_get_suspect_screening_csv_from_mol_files(
  files = mol_merck_2[grepl(example_key, mol_merck_2)],
  file = file.path("dev", "merck_peak_finding", "data", "ACC1_28203_suspects.csv")
)


get_suspects_screening_csv(
  suspects = data.table(
    name = "ACC1_28203",
    rt = 100,
    mol = mol_merck_2[grepl(example_key, mol_merck_2)]
  ),
  file = file.path("dev", "merck_peak_finding", "data", "ACC1_28203_suspects.csv")
)
```

# Top Down

List of targets (possibly semi generated by main structure, using expected fragmentation patterns as well as transformations, e.g., oxidation) are screened in the feature list and the SMILES of the targeted is evaluated agains the experimental MS2 spectra when available.

# Bottom Up (rcdk, cdk)

Formulae is generated for each feature using expected elements and the feature list is further reduced by excluding unlikely formulae. The remaining formulae are then used to generate possible structures which are further screened against the experimental MS2 spectra when available.
