---
title: "App Guide"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{App Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The StreamFind app is a graphical front-end for the same Project/Workflow/Method
framework that is available from R. It does not maintain a separate analysis
state. Instead, it reads and writes the same DuckDB-backed project data that the
R API uses.

# Start the app

The app can be launched in two ways.

## Start without a project

```{r, eval = FALSE}
library(StreamFind)
run_app()
```

This opens the home screen and lets you create or open a project interactively.

## Start from an existing project object

```{r, eval = FALSE}
nta <- open_ProjectNonTargetAnalysis(
  db = "streamfind.duckdb",
  project_id = "demo_nts"
)

nta$run_app()
```

This is usually the preferred route because the app starts with the active
project already resolved.

# Main areas of the app

The current application layout is organized around top-level tabs:

- `Home` for entry points and general package information
- `Project` for project metadata and project-level operations
- `Analyses` for imported file metadata and assignments
- `Explorer` for raw-data inspection
- `Workflow` for configuring and executing project-owned methods
- `Results` for domain-specific outputs such as features and suspects
- `Report` for report generation
- `Audit Trail` for reviewing execution history

Which controls appear inside these areas depends on the active project class.

# Typical usage pattern

## 1. Create or open a project

When working from the app alone, begin by selecting the project database, the
project class, and a `project_id`. If a matching project already exists, the app
reuses it. Otherwise, a new project entry is created.

## 2. Import analyses

For Mass Spec projects, import the analytical files into the project. The app
stores the analysis names together with replicate and blank assignments so they
can be reused by workflow steps and result plots.

## 3. Configure the workflow

In the `Workflow` tab, add `Method` steps supported by the active project class.
For `ProjectNonTargetAnalysis`, this includes native methods such as feature
finding, feature grouping, gap filling, blank subtraction, matrix suppression
correction, suspect screening, and MetFrag screening.

## 4. Run and monitor execution

Execute the workflow from the app. Progress messages, audit entries, and cached
results remain attached to the project, so the same project can be reopened
later from R or from another app session.

## 5. Inspect results

The `Results` and `Explorer` tabs expose project-specific tables and plots. For
non-target analysis projects this includes, for example:

- feature counts
- feature profiles
- TIC and chromatographic traces
- matrix suppression plots
- suspect tables and associated spectra views

# Relationship with the R API

The app and the R API are interchangeable interfaces over the same project
backend. A common pattern is:

```{r, eval = FALSE}
nta <- open_ProjectNonTargetAnalysis(
  db = "streamfind.duckdb",
  project_id = "demo_nts"
)

nta$run_workflow()
nta$run_app()
```

You can prepare a workflow in code, inspect the results in the app, and then
continue scripting against the same project afterward.

# Tips for project-backed use

- Keep one DuckDB file per study or per major analysis branch.
- Use stable `project_id` values to separate alternative workflows in the same
  database.
- Prefer `project$run_app()` when a project is already active in the R session.
- Reuse the app for exploration and reporting, not as a separate storage layer.

# Next articles

- `vignette("general_guide", package = "StreamFind")` introduces the project
  framework.
- `vignette("demo_nts", package = "StreamFind")` shows a full
  `ProjectNonTargetAnalysis` workflow.
