---
title: "Non-Target Analysis Demo"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Non-Target Analysis Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Goal

This article shows how the current `ProjectNonTargetAnalysis` framework can be
used to build, run, and inspect a non-target screening workflow. The example is
based on the package's project-backed architecture rather than the legacy
`MassSpecEngine` interface.

# Project setup

```{r, eval = FALSE}
library(StreamFind)

db_file <- "streamfind_nta.duckdb"
project_id <- "wastewater_demo"

ms_files <- c(
  "01_tof_ww_is_pos_blank-r001.mzML",
  "01_tof_ww_is_pos_blank-r002.mzML",
  "02_tof_ww_is_pos_influent-r001.mzML",
  "02_tof_ww_is_pos_influent-r002.mzML"
)

nta <- open_ProjectNonTargetAnalysis(
  db = db_file,
  project_id = project_id,
  file_paths = ms_files
)
```

The project persists imported files, metadata, workflow state, and downstream
results in a DuckDB database.

# Replicates and blanks

Replicate and blank assignments are part of the project state and are reused by
workflow steps such as blank subtraction and matrix suppression correction.

```{r, eval = FALSE}
nta$set_replicate_names(c(
  "blank-r001",
  "blank-r002",
  "influent-r001",
  "influent-r002"
))

nta$set_blank_names(c(
  "blank-r001",
  "blank-r002",
  "blank-r001",
  "blank-r002"
))
```

# Build the workflow

The current NTA workflow is defined from `Method_NonTargetAnalysis_*()`
constructors. A realistic example can look like this:

```{r, eval = FALSE}
workflow <- Workflow(list(
  Method_NonTargetAnalysis_FindFeatures(
    rtWindows = data.frame(rtmin = numeric(), rtmax = numeric()),
    ppmThreshold = 10,
    noiseThreshold = 250,
    minSNR = 3,
    minTraces = 3L,
    baselineWindow = 200,
    maxWidth = 250,
    baseQuantile = 0.99
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS1(
    filtered = FALSE,
    rtWindow = c(-1, 1),
    mzWindow = c(-1, 6),
    minTracesIntensity = 250,
    mzClust = 0.008,
    presence = 0.5
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS2(
    filtered = FALSE,
    minTracesIntensity = 10,
    isolationWindow = 1.3,
    mzClust = 0.008,
    presence = 0.5
  ),
  Method_NonTargetAnalysis_GroupFeatures(
    method = "none",
    rtDeviation = 5,
    ppm = 10,
    minSamples = 1L
  ),
  Method_NonTargetAnalysis_FillFeatures(
    withinReplicate = FALSE,
    filtered = FALSE,
    rtExpand = 10,
    mzExpand = 0.01,
    maxPeakWidth = 30
  ),
  Method_NonTargetAnalysis_CorrectMatrixSuppression(
    refBlankReplicate = NA_character_,
    mpRtWindow = 10
  ),
  Method_NonTargetAnalysis_BlankSubtraction(
    blankThreshold = 5,
    rtExpand = 10,
    mzExpand = 0.005
  ),
  Method_NonTargetAnalysis_FilterFeatures(
    minIntensity = 10000,
    removeIsotopes = TRUE,
    removeAdducts = TRUE,
    removeLosses = TRUE
  )
))
```

Attach the workflow to the project:

```{r, eval = FALSE}
set_workflow(nta, workflow)
```

# Run the workflow

```{r, eval = FALSE}
nta$run_workflow()
```

The workflow executes in order and writes results back into the project
database. Because the state is persistent, the same project can be reopened
later without rerunning previous steps.

# Inspect feature results

## Features table

```{r, eval = FALSE}
features_all <- get_features(nta)
features_filtered <- get_features(nta, filtered = TRUE)
```

Typical columns include feature identifiers, group membership, analysis name,
retention time, m/z, intensity, and correction factors written by matrix
suppression correction.

## Feature counts

```{r, eval = FALSE}
plot_features_count(nta, groupBy = "replicate")
```

## Feature profiles

```{r, eval = FALSE}
plot_features_profile(
  nta,
  groups = "FG1",
  groupBy = "replicate",
  corrected = TRUE
)
```

When `corrected = TRUE`, the intensity profile uses the feature-specific
`correction` factor computed by `Method_NonTargetAnalysis_CorrectMatrixSuppression()`.

# Inspect matrix suppression

The current framework exposes matrix suppression both as a table and as a plot.

```{r, eval = FALSE}
matrix_suppression <- get_matrix_suppression(
  nta,
  rtWindowVal = 10
)

plot_matrix_suppression(
  nta,
  rtWindowVal = 10,
  colorBy = "replicates"
)
```

The native implementation derives TIC-based suppression relative to associated
blank samples and can scale feature correction factors using surrounding
internal standards when these are available in the project.

# Add internal standards or suspect screening

The workflow can be extended with internal-standard detection and suspect
screening methods when reference data are available.

```{r, eval = FALSE}
workflow_extended <- Workflow(c(unclass(workflow), list(
  Method_NonTargetAnalysis_FindInternalStandards(
    suspects = internal_standards,
    ppm = 10,
    sec = 15,
    ppmMS2 = 10,
    mzrMS2 = 0.008,
    minCosineSimilarity = 0.7,
    minSharedFragments = 3L,
    filtered = TRUE
  ),
  Method_NonTargetAnalysis_SuspectScreening(
    suspects = suspects,
    ppm = 5,
    sec = 10,
    ppmMS2 = 10,
    mzrMS2 = 0.008,
    minCosineSimilarity = 0.7,
    minSharedFragments = 3L,
    filtered = TRUE
  )
)))
```

You can then retrieve suspect annotations with:

```{r, eval = FALSE}
suspects_found <- get_suspects(nta)
```

# Add MetFrag screening

MetFrag screening is also project-owned in the current framework:

```{r, eval = FALSE}
metfrag_step <- Method_NonTargetAnalysis_MetFragScreening(
  metfrag_path = "C:/Tools/MetFragCommandLine-2.5.0.jar",
  database_type = "LocalCSV",
  database_path = "transformation_products_template.csv",
  ppm = 10,
  sec = 15,
  ppmMS2 = 10,
  mzrMS2 = 0.008,
  top_n = 5L,
  score_types = "FragmenterScore",
  score_weights = 1,
  pre_processing_candidate_filter = c("UnconnectedCompoundFilter", "IsotopeFilter"),
  post_processing_candidate_filter = "InChIKeyFilter",
  maximum_tree_depth = 3L,
  number_threads = 1L,
  use_smiles = TRUE,
  filtered = FALSE,
  java_path = "java",
  run_dir = "",
  debug = FALSE
)
```

When `run_dir = ""`, StreamFind creates a timestamped folder under
`./log/metfrag/` and manages temporary MetFrag files there.

# Use the app for interactive review

Once the project has been populated from R, the same project can be opened in
the app for manual inspection:

```{r, eval = FALSE}
nta$run_app()
```

# Summary

The recommended NTA pattern in StreamFind is:

1. Open a `ProjectNonTargetAnalysis`.
2. Import analyses and assign replicates and blanks.
3. Define a validated `Workflow`.
4. Run the workflow into the project database.
5. Inspect features, matrix suppression, and suspects from R or the app.
