---
title: "Developer Guide"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Developer Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

This article summarizes the current StreamFind package architecture after the
transition from legacy analysis containers to persistent project classes. The
main design goal is that the R API, the native C++ backend, and the Shiny app
all operate on the same DuckDB-backed project state.

# Core architecture

## Project hierarchy

The public project hierarchy is:

- `Project`
- `ProjectMassSpec`
- `ProjectMassSpecSpectra`
- `ProjectMassSpecChromatograms`
- `ProjectNonTargetAnalysis`

`Project` owns the shared database connection, metadata, audit trail, cache, and
workflow serialization. Domain-specific project classes layer additional native
helpers and S3/R6 convenience methods on top of that base.

## Workflow model

Workflows are metadata containers:

- `Method` describes one processing step and its validated parameters
- `Workflow` is an ordered list of `Method` objects

Execution is delegated back to the owning project class through `run_method()`
and `run_workflow()`.

## Native backend

Project classes are backed by Rcpp-exposed native objects. The main native
implementation areas are:

- `src/project/` for shared project persistence and cache logic
- `src/mass_spec/` for shared Mass Spec import and raw-data helpers
- `src/nta/` for non-target analysis algorithms and project APIs

# Package layout

The current source layout is broadly:

- `R/class_Project*.R` for project R6 and S3 front-ends
- `R/class_Method*.R` and `R/class_Workflow.R` for workflow metadata
- `R/class_MethodsNonTargetAnalysis.R` for native NTA method constructors,
  validation, and `run.*` dispatch
- `R/fct_open_project_classes.R` for public `open_<ProjectClass>()` helpers
- `src/rcpp_project*_export.cpp` for Rcpp bridges
- `src/nta/*.cpp` and `src/nta/*.h` for native NTA implementations
- `R/app_*` for the Shiny app modules

# Public extension pattern

## Add a new project-owned method

For a new `ProjectNonTargetAnalysis` processing step, the usual path is:

1. Add a `Method_NonTargetAnalysis_<Name>()` constructor in
   `R/class_MethodsNonTargetAnalysis.R`.
2. Add `validate_object.Method_NonTargetAnalysis_<Name>()`.
3. Add `run.Method_NonTargetAnalysis_<Name>()` so the method can be executed by
   `run_method()` and `run_workflow()`.
4. Add or extend an Rcpp export in `src/rcpp_project_nta_export.cpp`.
5. Implement the native algorithm in `src/nta/`.
6. Expose the project-facing native entry point from `PROJECT_NON_TARGET_ANALYSIS`
   in `src/nta/nta.h` and `src/nta/nta.cpp`.

The method constructor should describe:

- the public method name
- required predecessor steps
- the owning project class
- parameter defaults
- developer and reference metadata

## Add a new project getter or plot helper

If the new capability is a retrieval or visualization helper rather than a
workflow step, the usual path is:

1. Add the native project method in C++ if data access is backend-owned.
2. Add the Rcpp wrapper.
3. Add the S3 function in the appropriate `R/class_Project*.R` file.
4. Add the matching R6 convenience wrapper when appropriate.
5. Document the new interface and, if relevant, surface it in the app.

# Persistence and caching

Unlike the older in-memory workflow style, the current framework stores project
state in DuckDB. This has several consequences:

- methods should write reusable outputs back into project-owned tables
- workflows should be serializable and reloadable
- expensive computations can use project cache entries
- app sessions and R sessions should see the same project state

For mutating NTA feature algorithms, the existing project methods frequently use
cached execution helpers before saving updated feature tables back to the
project.

# App integration

The app is no longer a separate orchestration layer. It consumes the same
project metadata and method registry as the R API. For developers, this means:

- new project classes should be discoverable through `projects_overview()`
- new project methods should have clean metadata and validation
- results should be readable through project getters so app modules can consume
  them directly

# Documentation pattern

When adding or changing functionality, keep these layers aligned:

- roxygen for R6 and S3 interfaces
- method constructor documentation
- relevant articles in `vignettes/articles/`
- `_pkgdown.yml` navigation and reference grouping when needed

# Minimal example: inspect the project registry

```{r}
StreamFind::projects_overview()
```

# Minimal example: inspect available NTA methods

```{r}
nta_methods <- StreamFind::projects_overview("ProjectNonTargetAnalysis")$processing_methods
head(nta_methods)
```

# Current development direction

The current framework emphasizes:

- project-backed reproducibility
- native algorithms with explicit method metadata
- a shared API surface between R and the app
- clearer separation between raw-data access, workflow execution, and result
  inspection

New development should generally extend this project-based model rather than the
legacy `Engine`/`Analyses` abstraction family.
