---
title: "General Guide"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{General Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

StreamFind is built around persistent DuckDB-backed project objects. Instead of
creating transient analysis containers in memory, you open a `Project` child
class, import analytical files, attach a `Workflow`, run project-owned
`Method`s, and then inspect or visualize the stored results either from R or
through the Shiny app.

The current public framework is centered on three concrete project classes:

- `ProjectMassSpecSpectra` for spectra import and raw spectra plots
- `ProjectMassSpecChromatograms` for chromatogram extraction and plots
- `ProjectNonTargetAnalysis` for feature-centric non-target screening workflows

# Core concepts

## `Project`

A `Project` is the persistent runtime and storage boundary. It owns:

- the DuckDB database file
- the active `project_id`
- imported analyses
- metadata and audit trail entries
- cached processing outputs
- the active `Workflow`

Projects can be reopened later from the same database and continue where a
previous R session stopped.

## `Workflow`

A `Workflow` is an ordered list of `Method` metadata objects. It describes which
processing steps should run, in which order, and with which parameter settings.

## `Method`

A `Method` is a validated processing-step descriptor owned by a project class.
For example, `Method_NonTargetAnalysis_FindFeatures()` configures the native
feature-finding routine for `ProjectNonTargetAnalysis`.

# Discover available project classes

The project registry can be inspected programmatically:

```{r}
StreamFind::projects_overview()
```

To inspect a single class:

```{r}
StreamFind::projects_overview("ProjectNonTargetAnalysis")
```

# Create or open a project

Most workflows start with one of the `open_<ProjectClass>()` helpers. These
functions either create a new project entry or reopen an existing one in the
same database.

```{r, eval = FALSE}
library(StreamFind)

db_file <- "streamfind.duckdb"

ms_files <- c(
  "sample_01.mzML",
  "sample_02.mzML",
  "blank_01.mzML"
)

nta <- open_ProjectNonTargetAnalysis(
  db = db_file,
  project_id = "demo_nts",
  file_paths = ms_files
)
```

You can also reopen the same project later without re-importing files:

```{r, eval = FALSE}
nta <- open_ProjectNonTargetAnalysis(
  db = "streamfind.duckdb",
  project_id = "demo_nts"
)
```

# Add analyses and experimental metadata

Mass spectrometry project classes share analysis-management helpers through
`ProjectMassSpec`.

```{r, eval = FALSE}
nta$get_analyses()
nta$get_analysis_names()

nta$set_replicate_names(c(
  "influent-r001",
  "influent-r002",
  "blank-r001"
))

nta$set_blank_names(c(
  "blank-r001",
  "blank-r001",
  "blank-r001"
))
```

Those assignments are stored in the project database and reused by downstream
methods such as blank subtraction, matrix suppression correction, and
replicate-level plots.

# Define a workflow

For non-target analysis, workflows are usually created from
`Method_NonTargetAnalysis_*()` constructors.

```{r, eval = FALSE}
workflow <- Workflow(list(
  Method_NonTargetAnalysis_FindFeatures(
    rtWindows = data.frame(rtmin = numeric(), rtmax = numeric()),
    ppmThreshold = 10,
    noiseThreshold = 250,
    minSNR = 3
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS1(
    filtered = FALSE,
    rtWindow = c(-1, 1),
    mzWindow = c(-1, 6)
  ),
  Method_NonTargetAnalysis_LoadFeaturesMS2(
    filtered = FALSE,
    isolationWindow = 1.3
  ),
  Method_NonTargetAnalysis_GroupFeatures(
    method = "none",
    rtDeviation = 5,
    ppm = 10
  ),
  Method_NonTargetAnalysis_FillFeatures(
    filtered = FALSE,
    rtExpand = 10,
    mzExpand = 0.01
  ),
  Method_NonTargetAnalysis_BlankSubtraction(
    blankThreshold = 5,
    rtExpand = 10,
    mzExpand = 0.005
  ),
  Method_NonTargetAnalysis_FilterFeatures(
    minIntensity = 10000,
    removeIsotopes = TRUE,
    removeAdducts = TRUE,
    removeLosses = TRUE
  )
))
```

Attach it to the project:

```{r, eval = FALSE}
nta$set_workflow(workflow)
```

The same workflow can be inspected later with:

```{r, eval = FALSE}
nta$get_workflow()
```

# Run the workflow

To execute the active workflow:

```{r, eval = FALSE}
nta$run_workflow()
```

A single step can also be run directly:

```{r, eval = FALSE}
step <- Method_NonTargetAnalysis_CorrectMatrixSuppression(
  refBlankReplicate = NA_character_,
  mpRtWindow = 10
)

nta$run_method(step)
```

Each method writes its results back into the shared project database and may
also populate the project cache for later reuse.

# Inspect results

Because all results are project-backed, post-processing helpers can be called at
any time after the relevant workflow steps have run.

```{r, eval = FALSE}
features <- get_features(nta, filtered = TRUE)
feature_counts <- nta$get_features_count(filtered = TRUE)
matrix_suppression <- get_matrix_suppression(nta)
suspects <- get_suspects(nta)
```

Common plotting helpers include:

```{r, eval = FALSE}
plot_spectra_tic(nta, groupBy = "replicate")
plot_features_count(nta, groupBy = "replicate")
plot_features_profile(nta, groups = "FG1", corrected = TRUE)
plot_matrix_suppression(nta, colorBy = "replicates")
```

# Use the app with the same project

The Shiny app uses the same project classes and database schema as the R API.
You can launch it with an already-open project:

```{r, eval = FALSE}
nta$run_app()
```

or without a project object:

```{r, eval = FALSE}
run_app()
```

Using `project$run_app()` is usually the easiest path during analysis because
the app starts with the active project context already loaded.

# Recommended workflow pattern

For reproducible work, the most practical pattern is:

1. Open a project in a dedicated DuckDB file.
2. Import analyses and define replicate and blank assignments.
3. Build a `Workflow` from validated `Method` constructors.
4. Run the workflow from R or the app.
5. Reopen the same project later for visualization, reporting, or extension.

# Next articles

- The app interface is described in `vignette("app_guide", package = "StreamFind")`.
- A complete non-target analysis example is shown in
  `vignette("demo_nts", package = "StreamFind")`.
- Package extension and developer patterns are summarized in
  `vignette("developer_guide", package = "StreamFind")`.
